---
title: "DrugDrug Interaction from BNF"
author: Dennis Wollersheim
output:
  workflowr::wflow_html
---

```{r loadlibrary,message=FALSE}

source('lib/function.R')

```

# methods
  1)  match BNF drugs to PBS drugs

#  Selection of interacting drugs from BNF file.

To begin with, the BNF file which was in pdf format, in which the names of the Drugs where coated in green colour, we first filtered and replaced those drugs name (i.e, Panadole -> ####Panadole%%%%) and converted the pdf file into text file.
After conversion, we wrote a regular expression script in Perl to only extract the drugs name. Each drugs were having interactions with some other multiple drugs. For example, 

Abacavir	-> RIBAVIRIN ; Abacavir	-> TIPRANAVIR; Abacavir -> ORLISTAT. Also Tipranavir -> CLARITHROMYCIN ; Tipranavir ->	RIFABUTIN ;Tipranavir	-> RIFAMPICIN.

The detail structure of pdf file to text file and further in our analysis is shown below:

```{r prepare_data}

read.table( 'data/drug_interactions.txt', sep='\t',header = FALSE) %>%
  as.tibble() %>% 
  mutate(V1=str_trim(str_to_upper(V1))) %>%
  mutate(V2=str_trim(str_to_upper(V2))) %>%
  rename(drug1=V1,drug2=V2) %>%
{ . } -> drug_drug_concordance

head(drug_drug_concordance)
```
# match BNF drugs to PBS drugs
Find all distinct drugs on both sides of interactions list

```{r}
drug_drug_concordance %>% 
  select(drug1) %>%
  bind_rows(drug_drug_concordance %>% select(drug2)) %>%
  distinct(drug2) %>%
  {.} -> final_drugs
```

# Matching the pbs items and atc code with the BNF drug drug interaction.

# Matching BNF only with PBS

```{r Selecting the pbs items and atc code to match with BNF data set}

my_db_get_query('Select * from pbs_item JOIN pbs_atc USING (atc_code)')%>%
  as.tibble()%>%
  mutate(drug_name = str_trim(drug_name)) %>%
  mutate(atc_meaning = str_trim(str_to_upper(atc_meaning))) %>%
  { .} ->pbs_drugs

write.csv(pbs_drugs,file="output/Spelling/MainList.csv")

cat("how many drugs from BNF have a direct match with DRUGS listed in PBS_ITEMS ")
final_drugs %>% 
  inner_join(pbs_drugs,by=c("drug2"="drug_name")) %>%
  distinct( drug2)%>%
  count()
```

```{r}
cat("how many drugs do not have as direct match")
final_drugs %>% 
  anti_join(pbs_drugs,by=c("drug2"="drug_name")) %>%
  count()
```

```{r}
final_drugs %>% 
  anti_join(pbs_drugs,by=c("drug2"="drug_name")) %>%
  rename( drug_name = drug2) %>%
  {.} -> drugs_to_match

head(drugs_to_match)
```
These are some drugs that we have to match either with ATC code or PBS_Items.

# TO Do
You need to match up the drugs_to_match with final_drugs.

  * Make a table for the many-many matching process
  * use the ( fuzzy_join package )[https://github.com/dgrtwo/fuzzyjoin] to try and match as much as you can.  
  * Please continue to do the work in this document
  * As compelling as it seems, please do not edit the drugdrug input file.  I want the work to be data driven as much as possible.

## groups 
  * possibly ending with S
  * you will need to do some research to find the groups

Numbers of drugs that needs to match are
```{r possible_drugs_to_match}
drugs_to_match %>% count()   
#259
```
The possible group of drug from BNF are as follows:
```{r}
drugs_to_match %>%
  filter( endsWith(drug_name, 'S')) %>%
  {.}->drug_groups

# 77 group of drugs; assumption made as name ending with S
head(drug_groups)
```

Out of 259 drugs which we need to match, 77 are groups, so remaining (181) are the one which we still need to look into for.
```{r}

drugs_to_match %>%
  filter( ! endsWith(drug_name, 'S')) %>%
  {.}->drug_not_matched

head(drug_not_matched)

```
# out of 259 not mathched drug, we have 77 group and 181 not matched drug
# Matching with ATC Code

```{r}
drug_not_matched %>%
  inner_join(pbs_drugs,by=c("drug_name"="atc_meaning")) %>%
  distinct(drug_name)%>%
  select(drug_name)%>%
  {.}->matched_non_group

cat("how many drugs FROM drugs_to_match (BNF) have a direct match to atc_meaning in main table(PBS)")
drug_not_matched %>%
  inner_join(pbs_drugs,by=c("drug_name"="atc_meaning")) %>%
  distinct(drug_name)%>%
  select(drug_name)%>%
  count()
```
Out of 181 not matched drug , we have 24 matching drug with atc_meaning, so now we need 181-24= 157 drug to match. Some of them are listed below:
```{r}
drug_not_matched %>%
  anti_join(pbs_drugs,by=c("drug_name"="atc_meaning")) %>%
  distinct(drug_name)%>%
  select(drug_name)%>%
  {.}->remaining_drug 

head(remaining_drug)
```

# Different method to verify result

##########
```{r}
drug_to_match_withoutNA <- na.omit(drugs_to_match)

main_table_withNA_ATC <- pbs_drugs%>%select(atc_meaning)

main_table_withNA_ATC %>%
  filter(atc_meaning!='NA') %>%
  distinct(atc_meaning)%>%
  {.} -> pure_ATC

pure_drug_maintable <- pbs_drugs %>% select(drug_name) %>% distinct(drug_name) 

pure_drug_pbs <- na.omit(pure_drug_maintable) %>% 
  rename(pbs_drug = drug_name)

cat("how many drugs FROM drugs_to_match have a direct match to atc_meaning in main table")
drug_not_matched %>%
  inner_join(pure_ATC,by=c("drug_name"="atc_meaning")) %>%
  count()
  
#########
```

# Matching and Spelling Checks for Drugs Not Matched
An assumption is made many drugs from not matched drugs list, there may be some spelling mistakes or naming error.
For instance, DIGOXIN in PBS may be written as DIGOXYIN in BNF or DIGOXIN in BNF may be similar to OXIN in PBS.
So to rectify these kinds of error we have done the following and listed some of those. And the detail can be found in output folder having Spelling_Check_Error folder.

Here we have checked for spelling difference between the drugs listed in concordance (BNF) and the main table pbs drug (PBS), and some of found mistake are listed below:
```{r}
remaining_drug$drug_name[1] <- "ST JOHNS WORT"

write.csv(remaining_drug,file="output/Spelling/TotalDrugsOtherThanGroupTOMatch.csv")

spelled_check <- remaining_drug %>% 
  stringdist_inner_join(pure_drug_pbs, by = c("drug_name" = "pbs_drug"),max_dist=3)

write.csv(spelled_check,file="output/Spelling/SpelledDrug.csv")

head(spelled_check)
```


Similarly, we have checked for spelling difference between the drugs listed in concordance (BNF) and the main table pbs drug (PBS) with ATC_Code, and some of found mistake are listed below:
```{r}

spelled_check_atc<- remaining_drug %>% 
  stringdist_inner_join(pure_ATC, by = c("drug_name" = "atc_meaning"),max_dist=3)

write.csv(spelled_check_atc,file="output/Spelling/SpelledDrug_ATC.csv")

head(spelled_check_atc)
```
```{r}
#Number of errored name drug matched with ATC code are: 46 
spelled_check_atc%>% 
  select(drug_name)%>% 
  distinct(drug_name)%>%count()

#Number of errored name drug matched with PBS ITEM are: 46 
spelled_check%>% 
  select(drug_name)%>% 
  distinct(drug_name)%>%count()
```

# Matching and Spelling Checks for Groups
```{r}
drug_groups %>%
  inner_join(pbs_drugs,by=c("drug_name"="atc_meaning")) %>%
  distinct(drug_name)%>%
  select(drug_name)%>%
  {.}->matched_group

cat("how many Group have a direct match to atc_meaning in main table")
drug_groups %>%
  inner_join(pbs_drugs,by=c("drug_name"="atc_meaning")) %>%
  distinct(drug_name)%>%
  select(drug_name)%>%
  count()
```

Out of 77 Group , we have 2 matching drug with atc_meaning, so now we need 77-2= 75 drug-groups to find. 
```{r}
drug_groups %>%
  anti_join(pbs_drugs,by=c("drug_name"="atc_meaning")) %>%
  distinct(drug_name)%>%
  select(drug_name)%>%
  {.}->remaining_groups

write.csv(remaining_groups,file="output/Spelling/TotalGroupsToMatch.csv")

# Remaining groups are those 75 whose spelling or name formation may be slightly different than they should be. 

head(remaining_groups)
```

Matching remaining drug-group with PBS_Items.
```{r}
groups_spelled_check <- remaining_groups %>% 
  stringdist_inner_join(pure_drug_pbs, by = c("drug_name" = "pbs_drug"),max_dist=3)

#count(groups_spelled_check) #13

write.csv(groups_spelled_check,file="output/Spelling/GroupSpelledDrug.csv")

head(groups_spelled_check)
```
Matching remaining drug-group with ATC_Code.
```{r}
Group_Spelled_check_atc<- remaining_groups %>% 
  stringdist_inner_join(pure_ATC, by = c("drug_name" = "atc_meaning"),max_dist=3)

#count(Group_Spelled_check_atc)

write.csv(Group_Spelled_check_atc,file="output/Spelling/Groups_SpelledDrug_ATC.csv")

head(Group_Spelled_check_atc)
```

```{r}
# Number of matching with ATC :10
Group_Spelled_check_atc%>% 
  select(drug_name)%>% 
  distinct(drug_name)%>%count()

# Number of matching with PBS : 9
groups_spelled_check%>% 
  select(drug_name)%>% 
  distinct(drug_name)%>%count()

```
# To Do for public health Students:

* Remaining groups or drugs, which we couldnot match either with ATC or PBS_Drug by any means are the jobs for the public health guys to figure it out.

```{r Mismatch drug for small dataset of patients i.e. pbs_rr done with 11 chunks}

#my_db_get_query("Select * from VW_PBS_INTERVAL_RR  where pin='00329081500' ")%>%
#  as.tibble()%>%
#  mutate(row=row_number()) %>%
#  select(pin,row,pbs_code,spply_dt,end_date) %>% 
#  rename( start=spply_dt ) %>%
#  rename (end=end_date) %>%
#{ .} ->pbs_interval_rr

#view(pbs_interval_rr)

####
#pbs_interval_rr %>%
#  interval_inner_join(pbs_interval_rr,by=c("start","end"),minoverlap = 7) %>%
#  select(pin.x,row.x,pbs_code.x,start.x,end.x,row.y,pbs_code.y,start.y,end.y)%>%
#  {.}-> joined_interval

small_fetch = FALSE

if(small_fetch==TRUE)
{
my_db_get_query("Select pin,pbs_code,spply_dt,pbs_rgltn24_adjst_qty qty from pbs_rr where pin like '00%' ")%>%
  as.tibble()%>%
  { . } -> rr_pbs_interval_00

my_db_get_query("Select pin,pbs_code,spply_dt,pbs_rgltn24_adjst_qty qty from pbs_rr where pin like '01%' ")%>%
  as.tibble()%>%
  { . } -> rr_pbs_interval_01

my_db_get_query("Select pin,pbs_code,spply_dt,pbs_rgltn24_adjst_qty qty from pbs_rr where pin like '02%' ")%>%
  as.tibble()%>%
  { . } -> rr_pbs_interval_02

my_db_get_query("Select pin,pbs_code,spply_dt,pbs_rgltn24_adjst_qty qty from pbs_rr where pin like '03%' ")%>%
  as.tibble()%>%
  { . } -> rr_pbs_interval_03

my_db_get_query("Select pin,pbs_code,spply_dt,pbs_rgltn24_adjst_qty qty from pbs_rr where pin like '04%' ")%>%
  as.tibble()%>%
  { . } -> rr_pbs_interval_04

my_db_get_query("Select pin,pbs_code,spply_dt,pbs_rgltn24_adjst_qty qty from pbs_rr where pin like '05%' ")%>%
  as.tibble()%>%
  { . } -> rr_pbs_interval_05

my_db_get_query("Select pin,pbs_code,spply_dt,pbs_rgltn24_adjst_qty qty from pbs_rr where pin like '06%' ")%>%
  as.tibble()%>%
  { . } -> rr_pbs_interval_06

my_db_get_query("Select pin,pbs_code,spply_dt,pbs_rgltn24_adjst_qty qty from pbs_rr where pin like '07%' ")%>%
  as.tibble()%>%
  { . } -> rr_pbs_interval_07

my_db_get_query("Select pin,pbs_code,spply_dt,pbs_rgltn24_adjst_qty qty from pbs_rr where pin like '08%' ")%>%
  as.tibble()%>%
  { . } -> rr_pbs_interval_08

my_db_get_query("Select pin,pbs_code,spply_dt,pbs_rgltn24_adjst_qty qty from pbs_rr where pin like '09%' ")%>%
  as.tibble()%>%
  { . } -> rr_pbs_interval_09

my_db_get_query("Select pin,pbs_code,spply_dt,pbs_rgltn24_adjst_qty qty from pbs_rr where pin like '1%' ")%>%
  as.tibble()%>%
  { . } -> rr_pbs_interval_1

rr_pin_var <- data.frame(ls(pattern = "rr_pbs_*"))

colnames(rr_pin_var) <- "pin"

head(rr_pin_var)

View(rr_pin_var)

cluster <- create_cluster(4)

for (pin in rr_pin_var[,]){
  nesting_function(pin,"rr")
}
}
```

```{r Mismatch drug for large dataset of patients i.e. pbs, done in 11 chunks}

Fetching_Big_data =FALSE

if(Fetching_Big_data==TRUE)
{
my_db_get_query("Select pin,pbs_code,spply_dt,pbs_rgltn24_adjst_qty qty from pbs where pin like '00%' ")%>%
  as.tibble()%>%
  { . } -> bigpbs_interval_00

my_db_get_query("Select pin,pbs_code,spply_dt,pbs_rgltn24_adjst_qty qty from pbs where pin like '01%' ")%>%
  as.tibble()%>%
  { . } -> bigpbs_interval_01

my_db_get_query("Select pin,pbs_code,spply_dt,pbs_rgltn24_adjst_qty qty from pbs where pin like '02%' ")%>%
  as.tibble()%>%
  { . } -> bigpbs_interval_02

my_db_get_query("Select pin,pbs_code,spply_dt,pbs_rgltn24_adjst_qty qty from pbs where pin like '03%' ")%>%
  as.tibble()%>%
  { . } -> bigpbs_interval_03

my_db_get_query("Select pin,pbs_code,spply_dt,pbs_rgltn24_adjst_qty qty from pbs where pin like '04%' ")%>%
  as.tibble()%>%
  { . } -> bigpbs_interval_04

my_db_get_query("Select pin,pbs_code,spply_dt,pbs_rgltn24_adjst_qty qty from pbs where pin like '05%' ")%>%
  as.tibble()%>%
  { . } -> bigpbs_interval_05

my_db_get_query("Select pin,pbs_code,spply_dt,pbs_rgltn24_adjst_qty qty from pbs where pin like '06%' ")%>%
  as.tibble()%>%
  { . } -> bigpbs_interval_06

my_db_get_query("Select pin,pbs_code,spply_dt,pbs_rgltn24_adjst_qty qty from pbs where pin like '07%' ")%>%
  as.tibble()%>%
  { . } -> bigpbs_interval_07

my_db_get_query("Select pin,pbs_code,spply_dt,pbs_rgltn24_adjst_qty qty from pbs where pin like '08%' ")%>%
  as.tibble()%>%
  { . } -> bigpbs_interval_08

my_db_get_query("Select pin,pbs_code,spply_dt,pbs_rgltn24_adjst_qty qty from pbs where pin like '09%' ")%>%
  as.tibble()%>%
  { . } -> bigpbs_interval_09

my_db_get_query("Select pin,pbs_code,spply_dt,pbs_rgltn24_adjst_qty qty from pbs where pin like '1%' ")%>%
  as.tibble()%>%
  { . } -> bigpbs_interval_1

write_csv <- FALSE

if (write_csv == TRUE) 
{
  write.csv(bigpbs_interval_00,file="data/BigChunks/pbs_interval_00.csv")
  
  write.csv(bigpbs_interval_01,file="data/BigChunks/pbs_interval_01.csv")
  
  write.csv(bigpbs_interval_02,file="data/BigChunks/pbs_interval_02.csv")
  
  write.csv(bigpbs_interval_03,file="data/BigChunks/pbs_interval_03.csv")
  
  write.csv(bigpbs_interval_04,file="data/BigChunks/pbs_interval_04.csv")
  
  write.csv(bigpbs_interval_05,file="data/BigChunks/pbs_interval_05.csv")
  
  write.csv(bigpbs_interval_06,file="data/BigChunks/pbs_interval_06.csv")
  
  write.csv(bigpbs_interval_07,file="data/BigChunks/pbs_interval_07.csv")
  
  write.csv(bigpbs_interval_08,file="data/BigChunks/pbs_interval_08.csv")
  
  write.csv(bigpbs_interval_09,file="data/BigChunks/pbs_interval_09.csv")
  
  write.csv(bigpbs_interval_1,file="data/BigChunks/pbs_interval_1.csv")
  
}
}

```




