---
title: "Who Drug Match"
author: "Nirmala"
date: "July 31, 2018"
output: html_document
---
```{r loadlibrary,message=FALSE}
 source('lib/function.R')  
```

# Match drugs from PBS to statins interaction paper
* Tasks:
1) read both statins and pbs_item data
2) create final statins_interactions.csv to store the final data that is cleaned


```{r prepare data}

 read.csv(file="data/hpy_interfer.csv",header=TRUE,sep=",") %>% 
  as.tibble() %>%
  rename( drug_name =Drugs, category=Category ) %>%
  mutate( drug_name  = str_to_upper( drug_name  ) ) %>% 
  { . } -> final_drugs


# Matching the pbs items and atc code with the BNF drug drug interaction.

# Matching BNF only with PBS

```{r Selecting the pbs items and atc code to match with BNF data set}

my_db_get_query('Select * from pbs_item JOIN pbs_atc USING (atc_code)')%>%
  as.tibble()%>%
  mutate(drug_name = str_trim(drug_name)) %>%
  mutate(atc_meaning = str_trim(str_to_upper(atc_meaning))) %>%
  { .} ->pbs_drugs


cat("how many drugs from BNF have a direct match with DRUGS listed in PBS_ITEMS ")

final_drugs %>% 
  inner_join(pbs_drugs,by=c("drug_name"="drug_name")) %>%
  distinct( drug_name)%>%
  count()

```

```{r}

cat("how many drugs do not have as direct match")
final_drugs %>% 
  anti_join(pbs_drugs) %>%
  count()

```

# list of drugs to match

```{r}

final_drugs %>%
  inner_join( pbs_drugs ) %>%
  distinct( category, drug_name ) %>% 
  { . } -> current_matches

final_drugs %>% 
  anti_join(pbs_drugs) %>%
  {.} -> drugs_to_match

drugs_to_match

```
These are some drugs that we have to match either with ATC code or PBS_Items.

Numbers of drugs that needs to match are

```{r possible_drugs_to_match}

drugs_to_match %>% count()   

drug_not_matched = drugs_to_match

```

# Matching drug_name with ATC meaning

```{r}

drug_not_matched %>%
  inner_join(pbs_drugs,by=c("drug_name"="atc_meaning")) %>%
  distinct(category, drug_name)%>%
  {.}->matched_via_drug_2_atc


matched_via_drug_2_atc %>%
  inner_join(pbs_drugs,by=c("drug_name"="atc_meaning")) %>%
  distinct(category, drug_name.y)%>%
  rename( drug_name = drug_name.y) %>%
  rbind( current_matches ) %>% 
  { . } -> current_matches

drug_not_matched %>%
  anti_join( matched_via_drug_2_atc ) %>%
  {.}-> drug_not_matched

```
## break up multiple drug_name 

```{r}

drug_not_matched %>%
  filter( !is.na( str_match( drug_name, '\\(I.E.,'))) %>%
  mutate( mcat = str_replace( drug_name, '.* \\(I.E., (.*)\\)','\\1')) %>%
  select(category, mcat) %>%
  separate( mcat, paste0('a',1:30), ',|/', fill="right") %>%
  gather( var, drug_name, -category) %>%
  mutate( drug_name = str_trim(drug_name )) %>%
  select( -var ) %>%
  distinct( category, drug_name ) %>% 
  { . } -> drugs_split


drugs_split %>%
  inner_join(pbs_drugs) %>%
  distinct( category, drug_name ) %>% 
  { . } -> matched_via_drug_split_2_drug_name 

drugs_split %>%
  inner_join(pbs_drugs,by=c("drug_name"="atc_meaning")) %>%
  distinct( category, drug_name ) %>% 
  { . } -> matched_via_drug_split_2_atc


matched_via_drug_split_2_drug_name %>%
  rbind( matched_via_drug_split_2_atc ) %>%
  rbind( current_matches ) %>% 
  { . } -> current_matches



drug_not_matched %>%
  rbind( drugs_split ) %>%
  arrange( category, drug_name )  %>%
  anti_join( matched_via_drug_split_2_drug_name ) %>%
  anti_join( matched_via_drug_split_2_atc ) %>% 
  filter( !is.na(drug_name )) %>%
  { . } -> drugs_to_find


drugs_to_find %>% write.csv( 'data/hpy_drugs_to_find.csv')
current_matches %>% write.csv( 'data/hpy_current_matches.csv')




```

# Matching and Spelling Checks for Drugs Not Matched
An assumption is made many drugs from not matched drugs list, there may be some spelling mistakes or naming error.
For instance, DIGOXIN in PBS may be written as DIGOXYIN in BNF or DIGOXIN in BNF may be similar to OXIN in PBS.
So to rectify these kinds of error we have done the following and listed some of those. And the detail can be found in output folder having Spelling_Check_Error folder.

Here we have checked for spelling difference between the drugs listed in concordance (BNF) and the main table pbs drug (PBS), and some of found mistake are listed below:

```{r}

pbs_drugs %>%
  distinct (drug_name ) %>%
  filter( !is.na(drug_name )) %>%
  stringdist_inner_join(drugs_to_find, by = c("drug_name" = "drug_name"),max_dist=3) %>% 
  rename( possible_misspelling = drug_name.x ) %>%
  rename( drug_name = drug_name.y ) %>%
  select( category, drug_name, possible_misspelling ) %>%
  right_join(drugs_to_find ) %>%
  mutate( possible_misspelling = ifelse( is.na(possible_misspelling ), '', possible_misspelling)) %>% 
  write.csv('data/hpy_to_match.csv')



pbs_drugs %>%
  distinct( drug_name, form_strength, atc_code, atc_meaning ) %>%
  filter( !is.na( atc_meaning ) & atc_meaning != 'NA') %>%
  group_by( drug_name, atc_code, atc_meaning ) %>%
  summarise( form_strength = paste( form_strength, collapse=', ')) %>%
  write.csv('data/hpy_pbs_item.csv')




```


