
---
title: "Updata Benzos"
author: "Dennis"
date: "September 26, 2018"
output: html_document
---
```{r loadlibrary,message=FALSE}

 library('tidyverse')
 library('workflowr')
 library('stringr')
 library(dplyr)
 source('lib/function.R')  
```

# Match Nirmala's current drug list

```{r prepare data}
pbs_who <- read.csv(file="data/pbs_who.csv",header=TRUE,sep=",") %>% 
  as.tibble() %>%
  rename( item_code = pbs_code,
         unit_wt = unit_wt_mg) %>%
  select( item_code, ddd_mg_factor, unit_wt, units )

benzo =read.csv(file="data/benzodiazepines.csv",header=TRUE,sep=",") %>% 
  as.tibble() %>%
  select (-ddd_mg_factor, -unit_wt, -units )


tibble( item_code = c("04522B", "04150K", "04151L", "04216X", "97997R"), 
       ddd_mg_factor = c( 7.5, 10, 10, 1, 10 ),
       unit_wt = c(7.5,3,6,1,10),
       units=replicate(5, 'mg' )
       )  %>% 
       { . } -> extras


benzo %>% 
  left_join( rbind( pbs_who, extras) , by='item_code' ) %>% 
  { . } -> final


library(RPostgreSQL)
drv <- dbDriver("PostgreSQL")
con <- dbConnect(drv, user='dewoller', dbname='mofi', host='localhost', password = Sys.getenv("PSSWD"))

table_name = "newDDD"
# Drop table if it already exists
if (dbExistsTable(con, table_name))
    dbRemoveTable(con, table_name)

# Write the data frame to the database
dbWriteTable(con, name = table_name, value = final, row.names = FALSE)


sql = '
UPDATE continuing.item b
SET ddd_mg_factor = a.ddd_mg_factor,
unit_wt = a.unit_wt,
units = a.units
FROM "newDDD" a
WHERE a.item_code = b.item_code;'
