---
title: "Who Drug Match"
author: "Nirmala"
date: "July 31, 2018"
output: html_document
---
```{r loadlibrary,message=FALSE}

library('tidyverse')
library('workflowr')
library('stringr')
library('dplyr')
library('readxl')
```

# Match drugs from Australian PBS to drugs from the WHO ddd drug list
* Tasks:
1) read both ddd_who and pbs_item data
2) create final pbs to store the final data that is cleaned

The main task is to match the drug form pbs_item with ddd_who table
* ddd_who: contains data from ddd_who table
* pbs_item: contains data from pbs-item table
* final_pbs: copy of pbs_item along with form_type, unit_wt and adm_route whose initial values are NA

```{r prepare data}
ddd_who <- read.csv(file="data/ddd_who.csv",header=TRUE,sep=";") %>%
  as.tibble()
pbs_item <- read.csv(file="data/pbs_item.csv",header=TRUE, sep=";")  %>%
  as.tibble()

#pbs_item %>% 
#  mutate( a=str_extract(drug_name,"mcg")) %>%
#  filter( !is.na(a))

#
#read.csv('/tmp/aitem.csv' ) %>% as.tibble() -> item
#
#read.csv('/tmp/continuing_item.csv' ) %>% as.tibble() -> citem
#
#item %>%
#  anti_join( citem, by=c('pbs_code'='item_code'))
#
#citem %>%
#  anti_join( item, by=c('item_code'='pbs_code')) %>% View


```

#create a copy of pbs_item dataframe and add three columns:form_type ,unit_wt,adm_route
```{r}

pbs_item %>% 
  mutate( form_type=NA,
            adm_route=NA,
            final_unit=NA,
            form_strength = tolower(final_pbs$form_strength)
            )  %>% 
  { . } -> final_pbs

```
## Categorise form_strengths
### Tasks:
1) find out what types (form_type)
2) find out how many grams  (unit_wt)
* words: contains the collection of possible categories that is figured out from form_strength of pbs_item
* x: list of the matching words in form_strength with words


```{r possible categories}
#list of types
words <- c("tablet","injection","eye drops" , "capsule","cream","ear drops","inhalation","mixtures","other","nasal","ointment","oral","transdermal","suppositories","mouth","lotions","intravesical","lemella","parenteral","i.v.","sublingual","rectal","implant","vaginal","lamella","syrup","elixir","lozenge","wafer", "pressurised inhalation", "powder for oral inhalation","solution for inhalation")

cat("What are the possible categories in form strength?")
words

# extract the list of matching words
x <- sapply(words, function(x) grepl(tolower(x), tolower(final_pbs$form_strength)))

# paste the matching words together
final_pbs$form_type <- apply(x, 1, function(i) paste0(names(i)[i], collapse = ","))

#final_pbs %>%
#  filter( str_detect(form_type, ',' )) 

```

#return all the row that have no specific form_type and replace by unknown
```{r} 

final_pbs$form_type <-final_pbs$form_type %>% replace(.=="","UNKNOWN")

```
# Extract the PBS administration route from the drug form_type field
## Determine adm_route according to the form_type
* use form_type to determine adm_route
* E.g. if form_type is Tablet then adm_route will be O i.e. Oral




# Dertermine the adm_route according to the form_type
* According to form_type determine the adm_route
* For eg: if form_type is Tablet then adm_route will be O i.e. Oral

```{r find admin route} 
final_pbs$adm_route[final_pbs$form_type=="oral"|final_pbs$form_type=="tablet"|final_pbs$form_type=="capsule"|final_pbs$form_type=="mouth"|final_pbs$form_type == "lozenge"|final_pbs$form_type == "elixir"|final_pbs$form_type == "wafer"|final_pbs$form_type == "syrup"]="O"
final_pbs$adm_route[final_pbs$form_type=="nasal"]="N"
final_pbs$adm_route[final_pbs$form_type=="ointment"]="Ointment"
final_pbs$adm_route[final_pbs$form_type=="instillation"]="Instill"
final_pbs$adm_route[final_pbs$form_type=="cream"|final_pbs$form_type=="lotions"|final_pbs$form_type=="transdermal"]="TD"
final_pbs$adm_route[final_pbs$form_type=="lamella"]="lamella"
final_pbs$adm_route[final_pbs$form_type=="rectal"]="R"
final_pbs$adm_route[final_pbs$form_type=="vaginal"]="V"
final_pbs$adm_route[final_pbs$form_type=="injection"|final_pbs$form_type=="eye lotion"| final_pbs$form_type =="eye drops"|final_pbs$form_type=="ear drops"|final_pbs$form_type=="parenteral"|final_pbs$form_type=="i.v."]="P"
final_pbs$adm_route[final_pbs$form_type=="inhalation"]="Inhal"
final_pbs$adm_route[final_pbs$form_type=="implant"]="Implant"
final_pbs$adm_route[final_pbs$form_type=="intravesical"] = "Intravesical"
final_pbs$adm_route[final_pbs$form_type=="UNKNOWN"]="UNKNOWN"
final_pbs$adm_route[final_pbs$form_type=="sublingual"]="SL"
final_pbs$adm_route[final_pbs$form_strength == "powder for oral inhalation"] = "Inhal.powder"
final_pbs$adm_route[final_pbs$form_strength == "solution for inhalation"] = "Inhal.sol"
final_pbs$adm_route[final_pbs$form_strength == "pressurised inhalation"] = "Inhal.aerosol"

```


# Determine mg dosage of active ingredient of each PBS drug
## Tasks
1) change micrograms into mg 
2) get the amount
## Notes: 
* Any micrograms in form_strength is converted to mg so that it is easy to get any number before mg
* Using regular expression find the number of mg and enter it in unit_wt column of final_pbs table

```{r} 
final_pbs$form_strength <- str_replace(final_pbs$form_strength, "micrograms", "mg")
final_pbs$form_strength <- str_replace(final_pbs$form_strength, "microgram", "mg")

#get unit_wt 
final_pbs$unit_wt_mg <- str_extract(final_pbs$form_strength, "\\d*\\.*\\d+\\s?mg") 
final_pbs$unit_wt_g <- str_extract(final_pbs$form_strength, "\\d*\\.*\\d+\\s?g")
final_pbs$unit_wt_ml <-str_extract(final_pbs$form_strength,"per .*\\d*\\s?ml")


#remove g/mg from each unit_wt_g and unit_wt_mg column
final_pbs$unit_wt_mg  <- str_replace(final_pbs$unit_wt_mg , '\\mg', '')
final_pbs$unit_wt_g  <- str_replace(final_pbs$unit_wt_g , '\\g', '')
final_pbs$unit_wt_g<- as.numeric(final_pbs$unit_wt_g)
final_pbs$unit_wt_mg<- as.numeric(final_pbs$unit_wt_mg)
final_pbs$final_unit<- as.numeric(final_pbs$final_unit)

#get the final unit

final_pbs %>%
  mutate( final_unit = 
         ifelse( !is.na(unit_wt_g) & !is.na(unit_wt_mg),
                unit_wt_mg+unit_wt_g*1000, 
                ifelse( is.na(unit_wt_g), 
                    unit_wt_mg,
                    ifelse( is.na(unit_wt_mg),
                      unit_wt_g*1000,
                      final_unit )))) %>% 
                      { . } -> final_pbs 

#filter( !is.na( final_unit )) %>%   
#filter( pbs_code == '08449Q')  %>%

```
# clean up DDD administration route for better matching
Clean adm_route of ddd_who to match with pbs_item

Change the adm_route and make it equivalent to form_type of final_pbs table so that it will be easy to match E.g. Inh.sol,Inl.aersol to Inhal
          

```{r} 
ddd_who$adm_route <- str_replace(ddd_who$adm_route, "TD.*_*", "TD")
ddd_who$adm_route <- str_replace(ddd_who$adm_route, "oral.*_*", "O")

```

# match up the DDD and PBS, using the atc codes and the cleaned administration routes

Join both tables using atc code and admin route; 

## Tasks
1) get distinct pbs_code from existing pbs table
2) write in to a disk file for further processing

To narrow down the search, we filter out pbs_item drugs without corresponding DDD atc_code 

  * existing_pbs :list of distinct pbs_code  from pbs
  * existing_pbs_to match: those list of drug that is required to be match with ddd_who 
  * unknown_routes: those list of drug that needs the specific adm routes

 
#get distinct pbs_code for first time
#my_db_get_query("select distinct pbs_code from pbs") -> existing_pbs
#write.csv(existing_pbs, file = 'data/existing_pbs.csv')
```{r}
#get existing pbs data from csv file
existing_pbs <- data.frame(read.csv(file="data/existing_pbs.csv",header=TRUE, sep=",")[,2]) 
colnames(existing_pbs)<- 'pbs_code'
#filter those drugs that only have atc code in ddd_who table to narrow down search
final_pbs%>%
  inner_join(existing_pbs, by = c("pbs_code"="pbs_code"))%>%
  filter(atc_code %in% ddd_who$atc)%>%
  mutate(uadm_route = adm_route)%>%
  {.} ->existing_pbs_to_match


cat("How many specific adm routes?")
existing_pbs_to_match %>% count(adm_route,sort = TRUE)


```
#find the unknown routes
```{r}
#now 159 unknown
existing_pbs_to_match%>%
  filter(is.na(existing_pbs_to_match$adm_route)|existing_pbs_to_match$adm_route=="UNKNOWN")%>%
  {.} ->unknown_routes

write.csv(unknown_routes, file = 'data/total_unknown_routes.csv')
```

#existing pbs to match with combing total unknown routes  2689
#read data from Sedar
```{r}

read_excel('output/total_unknown_routes.xlsx')%>%
  select(-1)%>%
  mutate(final_unit=as.numeric(final_unit),
         unit_wt_mg = as.numeric(unit_wt_mg),
         unit_wt_g = as.numeric(unit_wt_g))%>%
 # filter(is.na(unit_wt_mg))%>%
  {.} -> figured_out_routes

```
#first remove and data with na and unknown from existing_pbs_to_match then
#bind the data
```{r}

existing_pbs_to_match %>% 
  anti_join(unknown_routes, by=c('pbs_code'))%>%
  bind_rows(figured_out_routes)%>%
  {.} ->final_existing_pbs_to_match

```



```{r}

#find the unknown final_unit in existing_pbs_to_match
existing_pbs_to_match%>% filter(!is.na(existing_pbs_to_match$unit_wt_ml)|is.na(existing_pbs_to_match$final_unit))%>%
{.} ->unknown_quantity

#336 unknown quantity
write.csv(unknown_quantity, file = 'data/unknown_quantity.csv')

```

#find the unkown units and combine to final_pbs_to_match

```{r}

read_excel('output/unknown_quantity.xlsx')%>%
  select(-1)%>%
  mutate(final_unit=as.numeric(final_unit),
         unit_wt_mg = as.numeric(unit_wt_mg),
         unit_wt_g = as.numeric(unit_wt_g))%>%
 # filter(is.na(unit_wt_mg))%>%
  {.} -> figured_out_quantity

```

#Add a coulmn units in final_existing_pbs_to_match and rename final_unit to quantity
```{r}

colnames(final_existing_pbs_to_match)[which(names(final_existing_pbs_to_match) == "final_unit")] <- "quantity"
final_existing_pbs_to_match$units = "mg"

```
#bind the data
```{r}

final_existing_pbs_to_match %>% 
  anti_join(unknown_quantity, by=c('pbs_code'))%>%
  bind_rows(figured_out_quantity)%>%
  {.} ->all_existing_pbs_to_match
all_existing_pbs_to_match%>%filter(is.na(quantity))
all_existing_pbs_to_match%>%filter(is.na(adm_route))
all_existing_pbs_to_match%>%filter(adm_route=="UNKNOWN")

```


# TO Do
* find all the routes for unknown routes in unknown_routes.csv
* finally join all_existing_pbs_to _match with ddd_who

```{r match with ddd_who table, warning=FALSE}

all_existing_pbs_to_match%>%
  inner_join(ddd_who, c("atc_code" = "atc","adm_route" = "adm_route")) %>%
  { .} ->pbs_who

```
