---
title: "Who Drug Match"
author: "Nirmala"
date: "July 31, 2018"
output: html_document
---
```{r loadlibrary,message=FALSE}

 library('tidyverse')
 library('workflowr')
 library('stringr')
 library(dplyr)
 source('lib/function.R')  
```

# Match drugs from Australian PBS to drugs from the WHO ddd drug list
* Tasks:
1) read both ddd_who and pbs_item data
2) create final pbs to store the final data that is cleaned

The main task is to match the drug form pbs_item with ddd_who table
* ddd_who: contains data from ddd_who table
* pbs_item: contains data from pbs-item table
* final_pbs: copy of pbs_item along with form_type, unit_wt and adm_route whose initial values are NA

```{r prepare data}
ddd_who <- read.csv(file="data/ddd_who.csv",header=TRUE,sep=";") %>% as.tibble()
pbs_item <- read.csv(file="data/pbs_item.csv",header=TRUE, sep=";")  %>% as.tibble()


pbs_item %>% 
  mutate( a=str_extract(drug_name,"mcg")) %>%
  filter( !is.na(a))


read.csv('/tmp/aitem.csv' ) %>% as.tibble() -> item

read.csv('/tmp/continuing_item.csv' ) %>% as.tibble() -> citem

item %>%
  anti_join( citem, by=c('pbs_code'='item_code'))

citem %>%
  anti_join( item, by=c('item_code'='pbs_code')) %>% View



#create a copy of pbs_item dataframe and add three columns:form_type ,unit_wt,adm_route
final_pbs <- pbs_item
final_pbs["form_type"] <- NA
final_pbs["adm_route"] <-NA
final_pbs["final_unit"] <- NA
final_pbs["form_strength"] <-tolower(final_pbs$form_strength)
```


# Try to categorise the different form_strengths
* Tasks:
1) find out what types they are
2) find out how many grams they are

## Categorise form_strengths
### Tasks:
1) find out what types (form_type)
2) find out how many grams  (unit_wt)



* words: contains the collection of possible categories that is figured out from form_strength of pbs_item
* x: list of the matching words in form_strength with words


```{r possible categories}
#list of types
words <- c("tablet","injection","eye drops" , "capsule","cream","ear drops","inhalation","mixtures","other","nasal","ointment","oral","transdermal","suppositories","mouth","lotions","intravesical","lemella","parenteral","i.v.","sublingual","rectal","implant","vaginal","lamella","syrup","elixir","lozenge","wafer", "pressurised inhalation", "powder for oral inhalation","solution for inhalation")

cat("What are the possible categories in form strength?")
words

# extract the list of matching words
x <- sapply(words, function(x) grepl(tolower(x), tolower(final_pbs$form_strength)))

# paste the matching words together
final_pbs$form_type <- apply(x, 1, function(i) paste0(names(i)[i], collapse = ","))

```



Powder,aerosal,solution == inhalation
```{r} 


#final_pbs%>% distinct(form_strength) %>% filter(!is.na(str_match(form_strength, ".*inh.*")))

#return all the row that have no specific form_type and replace by unknown

final_pbs$form_type <-final_pbs$form_type %>% replace(.=="","UNKNOWN")





```
# Extract the PBS administration route from the drug form_type field
## Determine adm_route according to the form_type
* use form_type to determine adm_route
* E.g. if form_type is Tablet then adm_route will be O i.e. Oral




# Dertermine the adm_route according to the form_type
* According to form_type determine the adm_route
* For eg: if form_type is Tablet then adm_route will be O i.e. Oral

```{r find admin route} 
final_pbs$adm_route[final_pbs$form_type=="oral"|final_pbs$form_type=="tablet"|final_pbs$form_type=="capsule"|final_pbs$form_type=="mouth"|final_pbs$form_type == "lozenge"|final_pbs$form_type == "elixir"|final_pbs$form_type == "wafer"|final_pbs$form_type == "syrup"]="O"
final_pbs$adm_route[final_pbs$form_type=="nasal"]="N"
final_pbs$adm_route[final_pbs$form_type=="ointment"]="Ointment"
final_pbs$adm_route[final_pbs$form_type=="instillation"]="Instill"
final_pbs$adm_route[final_pbs$form_type=="cream"|final_pbs$form_type=="lotions"|final_pbs$form_type=="transdermal"]="TD"
final_pbs$adm_route[final_pbs$form_type=="lamella"]="lamella"
final_pbs$adm_route[final_pbs$form_type=="rectal"]="R"
final_pbs$adm_route[final_pbs$form_type=="vaginal"]="V"
final_pbs$adm_route[final_pbs$form_type=="injection"|final_pbs$form_type=="eye lotion"| final_pbs$form_type =="eye drops"|final_pbs$form_type=="ear drops"|final_pbs$form_type=="parenteral"|final_pbs$form_type=="i.v."]="P"
final_pbs$adm_route[final_pbs$form_type=="inhalation"]="Inhal"
final_pbs$adm_route[final_pbs$form_type=="implant"]="Implant"
final_pbs$adm_route[final_pbs$form_type=="intravesical"] = "Intravesical"
final_pbs$adm_route[final_pbs$form_type=="UNKNOWN"]="UNKNOWN"
final_pbs$adm_route[final_pbs$form_type=="sublingual"]="SL"
final_pbs$adm_route[final_pbs$form_strength == "powder for oral inhalation"] = "Inhal.powder"
final_pbs$adm_route[final_pbs$form_strength == "solution for inhalation"] = "Inhal.sol"
final_pbs$adm_route[final_pbs$form_strength == "pressurised inhalation"] = "Inhal.aerosol"



```


# Determine mg dosage of active ingredient of each PBS drug
## Tasks
1) change micrograms into mg 
2) get the amount
## Notes: 
* Any micrograms in form_strength is converted to mg so that it is easy to get any number before mg
* Using regular expression find the number of mg and enter it in unit_wt column of final_pbs table

```{r} 
final_pbs$form_strength <- str_replace(final_pbs$form_strength, "micrograms", "mg")
final_pbs$form_strength <- str_replace(final_pbs$form_strength, "microgram", "mg")
#get unit_wt 

#final_pbs$unit_wt_mg <- str_extract(final_pbs$form_strength, "\\d+\\.*\\d+\\s?mg") 
final_pbs$unit_wt_mg <- str_extract(final_pbs$form_strength, "\\d*\\.*\\d+\\s?mg") 
#final_pbs$unit_wt_g <- str_extract(final_pbs$form_strength, "\\d+\\s?g") 
final_pbs$unit_wt_g <- str_extract(final_pbs$form_strength, "\\d*\\.*\\d+\\s?g")
final_pbs$unit_wt_ml <-str_extract(final_pbs$form_strength,"per .*\\d*\\s?ml")


#remove g/mg from each unit_wt_g and unit_wt_mg column
final_pbs$unit_wt_mg  <- str_replace(final_pbs$unit_wt_mg , '\\mg', '')
final_pbs$unit_wt_g  <- str_replace(final_pbs$unit_wt_g , '\\g', '')


final_pbs$unit_wt_g<- as.numeric(final_pbs$unit_wt_g)
final_pbs$unit_wt_mg<- as.numeric(final_pbs$unit_wt_mg)
final_pbs$final_unit<- as.numeric(final_pbs$final_unit)

#get the final unit

for (i in 1:7614)
  if(!is.na(final_pbs$unit_wt_g[i]) & !is.na(final_pbs$unit_wt_mg[i])){
      final_pbs$final_unit[i] =final_pbs$unit_wt_mg[i]+final_pbs$unit_wt_g[i]*1000
    }else{
      if(is.na(final_pbs$unit_wt_g[i])){
          final_pbs$final_unit[i] =final_pbs$unit_wt_mg[i]
        }else if(is.na(final_pbs$unit_wt_mg[i])){
            final_pbs$final_unit[i] = final_pbs$unit_wt_g[i]*1000
    }
}








final_pbs %>% distinct(final_pbs$unit_wt_ml)
```
# clean up DDD administration route for better matching
Clean adm_route of ddd_who to match with pbs_item

Change the adm_route and make it equivalent to form_type of final_pbs table so that it will be easy to match E.g. Inh.sol,Inl.aersol to Inhal
          

```{r} 
#ddd_who$adm_route <- str_replace(ddd_who$adm_route, "Inh.*_*", "Inhal")
ddd_who$adm_route <- str_replace(ddd_who$adm_route, "TD.*_*", "TD")
ddd_who$adm_route <- str_replace(ddd_who$adm_route, "oral.*_*", "O")

```

# match up the DDD and PBS, using the atc codes and the cleaned administration routes

Join both tables using atc code and admin route; 

## Tasks
1) get distinct pbs_code from existing pbs table
2) write in to a disk file for further processing

To narrow down the search, we filter out pbs_item drugs without corresponding DDD atc_code 

  * existing_pbs :list of distinct pbs_code  from pbs
  * existing_pbs_to match: those list of drug that is required to be match with ddd_who 
  * unknown_routes: those list of drug that needs the specific adm routes

 
#get distinct pbs_code for first time
#my_db_get_query("select distinct pbs_code from pbs") -> existing_pbs
#write.csv(existing_pbs, file = 'data/existing_pbs.csv')

 ```{r  }

#get existing pbs data from csv file
existing_pbs <- read.csv(file="data/existing_pbs.csv",header=TRUE, sep=",")[,2] %>%
  as.tibble()

colnames(existing_pbs)<- 'pbs_code'



#filter those drugs that only have atc code in ddd_who table to narrow down search
final_pbs%>%
  inner_join(existing_pbs, by = c("pbs_code"="pbs_code"))%>%
  filter(atc_code %in% ddd_who$atc)%>%
  mutate(uadm_route = adm_route)%>%
  {.} ->existing_pbs_to_match


cat("How many specific adm routes?")
existing_pbs_to_match %>% count(adm_route,sort = TRUE)


#total 82 unknown routes 
#now 159 unknown
existing_pbs_to_match%>%
  filter(is.na(existing_pbs_to_match$adm_route)|existing_pbs_to_match$adm_route=="UNKNOWN")%>%
  {.} ->unknown_routes

write.csv(unknown_routes, file = 'data/total_unknown_routes.csv')


unknown_routes_old <- read.csv(file="data/unknown_routes.csv",header=TRUE,sep=",")

unknown_routes_old<-subset(unknown_routes_old,select = -c(X))


#unknown_routes_old["final_unit"] <-NA
#unknown_routes_old["unit_wt_mg"] <-NA 
#unknown_routes_old["unit_wt_g"] <- NA
#unknown_routes_old["unit_wt_ml"] <-NA

#unknown_routes_old["final_unit"] <-as.numeric(unknown_routes_old["final_unit"])
#unknown_routes_old["unit_wt_mg"] <- as.numeric(unknown_routes_old["unit_wt_mg"])
#unknown_routes_old["unit_wt_g"] <- as.numeric(unknown_routes_old["unit_wt_g"])
#unknown_routes_old["unit_wt_ml"] <- as.character(unknown_routes_old["unit_wt_ml"])


#total 77 obs.
new_unknown_routes<-unknown_routes[!unknown_routes$pbs_code%in%unknown_routes_old$pbs_code,]



write.csv(unknown_routes, file = 'data/new_unknown_routes.csv')



#find the unknown final_unit in existing_pbs_to_match
existing_pbs_to_match %>% 
  filter(!is.na(existing_pbs_to_match$unit_wt_ml) | is.na(existing_pbs_to_match$final_unit)) %>%
  {.} ->unknown_quantity



#336 unknown quantity
write.csv(unknown_quantity, file = 'data/unknown_quantity.csv')

#update the adm_routes having NA values:
#existing_pbs_to_match$uadm_route[existing_pbs_to_match$atc_code == "R03AC02"] <- "Inhal"
```


# TO Do
* find all the routes for unknown routes in unknown_routes.csv
* finally join existing_pbs_to _match with ddd_who
```{r match with ddd_who table, warning=FALSE}

existing_pbs_to_match%>%
  inner_join(ddd_who, c("atc_code" = "atc","adm_route" = "adm_route")) %>%
  { .} ->pbs_who


pbs_who %>% count(adm_route,sort = TRUE)

pbs_who %>% write.csv( 'data/pbs_who.csv', row.names=FALSE )



```
